Code Clones: To refactor or not to refactor?

### 2020-05-09 ase/paper.tex
Updates from Overleaf

In this study, we analyze the effect of clone refactoring on maintainability. We create a tool that automatically applies ``Extract Method'' refactorings to a subset of detected clones. We measure several maintainability metrics \cite{heitlager2007practical} before and after refactoring each clone to determine its impact on system maintainability. These metrics are volume, number of parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of projects in a controlled environment. We use this data to understand which characteristics of clones have the biggest influence on system maintainability when refactoring. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quantile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.

@@

This paper is organized as follows. In Sec.~\ref{sec:background} we revisit key clone refactoring background. Sec.~\ref{sec:relatedwork} discusses state-of-the-art in automatic clone refactoring research. In Sec.~\ref{sec:research} we propose a definition for clones such that they could and should be automatically refactored. In Sec.~\ref{sec:clonerefactor} we introduce our clone refactoring tool. In Sec.~\ref{sec:experimentalsetup} we explain the setup of our experiments; we show the results in Sec.~\ref{sec:results} and discuss them in Sec.~\ref{sec:discussion}.
%We evaluate our approach by comparing the maintainability of before- and after snapshots of a codebase for each duplication problem that is refactored. We define which factors have the greatest influence on whether a clone improves maintainability when refactored. We find a set of thresholds by which clones can be detected that are more likely to improve maintainability when refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

A higher type of clone means that it is harder to detect and refactor. Many studies adopt these clone types, analyzing them further and writing detection techniques for them \cite{sajnani2016sourcerercc, kodhai2010detection, van2019novel}.

\subsection{Clone harmfulness}
\todo{incomplete}

@@

\subsection{Clone context analysis}\label{sec:rw:contextanalysis}
Golomingi~\cite{koni2001scenario} explores mapping the relation between clone instances to refactoring methods. The author analyses the refactoring methods described by Martin Fowler \cite{fowler1999refactoring} and analyzes what refactoring methods can be used to refactor clones with what inheritance relations. The identified clone relations are: Ancestor, Common Hierarchy, First Cousin, Same Method, Sibling, Single Class, Superclass and Unrelated. We extend this list with several more fine-grained relations, suitable for automatic refactoring (see Sec.~\ref{sec:relation}).

@@

\section{Defining refactorable clones}\label{sec:research}
%In literature, several clone type definitions have been used to argue about duplication in source code~\cite{roy2007survey}.
We discuss how we can define clones such that they can be automatically refactored while avoiding changing the functional behavior and other negative side effects on the source code, such as decreasing the maintainability of the code.

>>>>

In this study, we analyze which factors influence the impact of refactoring a clone on maintainability. We create a tool that automatically applies ``Extract Method'' refactorings to a subset of detected clones. We measure several maintainability metrics \cite{heitlager2007practical} before and after refactoring each clone to determine its impact on system maintainability. These metrics are volume, number of parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of projects in a controlled environment. We use this data to understand which characteristics of clones have the biggest influence on system maintainability when refactoring. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quantile distribution of empirical data~\cite{alves2010deriving}. Using the delta maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.

@@

This paper is organized as follows. In Sec.~\ref{sec:background} we revisit key clone refactoring background. Sec.~\ref{sec:relatedwork} discusses state-of-the-art in automatic clone refactoring research. In Sec.~\ref{sec:research} we propose preconditions for a subset of clones such that they can be automatically refactored. In Sec.~\ref{sec:clonerefactor} we introduce our clone refactoring tool. In Sec.~\ref{sec:experimentalsetup} we explain the setup of our experiments; we show the results in Sec.~\ref{sec:results} and discuss them in Sec.~\ref{sec:discussion}.
%We evaluate our approach by comparing the maintainability of before- and after snapshots of a codebase for each duplication problem that is refactored. We define which factors have the greatest influence on whether a clone improves maintainability when refactored. We find a set of thresholds by which clones can be detected that are more likely to improve maintainability when refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

A higher type of clone means that it is harder to detect and refactor. Many studies adopt these clone types, analyzing them further and writing detection techniques for them~\cite{sajnani2016sourcerercc, kodhai2010detection, van2019novel}.

\subsection{Clone harmfulness}
In literature, clones are generally considered anti-patterns that should be avoided~\cite{fowler2018refactoring, ostberg2014automatically, chatterji2013effects}. However, removing some clones can leave the system in a less-desirable state~\cite{jarzabek2010clones, kapser2008cloning}. Jarzabek et al.~\cite{jarzabek2010clones} describe that mainly clones located in semantically unrelated parts of the codebase hinder maintenance. Clones in semantically unrelated parts of the codebase are merely a manifestation of the design choices. Therefore, elimination of clones is only beneficial if it also reduces
the semantic dependency among parts of the system.

@@

\subsection{Clone context analysis}\label{sec:rw:contextanalysis}
Golomingi~\cite{koni2001scenario} explores mapping the relation between clone instances to refactoring methods. The author analyses the refactoring methods described by Martin Fowler~\cite{fowler1999refactoring} and analyzes what refactoring methods can be used to refactor clones with what inheritance relations. The identified clone relations are: Ancestor, Common Hierarchy, First Cousin, Same Method, Sibling, Single Class, Superclass and Unrelated. We extend this list with several more fine-grained relations, suitable for automatic refactoring (see Sec.~\ref{sec:relation}).

@@

\section{Defining automatically refactorable clones}\label{sec:research}
%In literature, several clone type definitions have been used to argue about duplication in source code~\cite{roy2007survey}.
We discuss how we can define clones such that they can be automatically refactored while avoiding changing the functional behavior and other negative side effects on the source code, such as decreasing the maintainability of the code.
